---
title: "ElasticNet Regression Part I"
author: "Jeffrey Strickland"
date: '2022-08-28'
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, dpi = 300)
```



## Data Preprocessing
To start, we’ll load the data from the data we saved earlier.

## Load Libraries

```{r}
# ggplot for system for 'declaratively' creating graphics [2]
library(ggplot2) 
# dplyrfor the grammar of data manipulation (i.e., %>%, the pipe operator)
library(dplyr) 
# superml provides a scikit-learn's fit, predict interface to train machine learning models in R
library(superml) 
# readr for reading and writing files (.csv) in R
library(readr)  
# glmnet for fitting generalized linear models via penalized maximum likelihood.
library(glmnet)
# plyr for the split-apply-combine paradigm for R
library(plyr)
# for miscellaneous functions for training and plotting classification and regression models
library(caret) 
# Metrics provides evaluation metrics in R that are commonly used in supervised machine learning.
library(Metrics)
# rsq for generalized R-squared, partial R-squared, and partial correlation coefficients for generalized linear models and linear (mixed) models
library(rsq)
# knitr for R markdown and fancy output like tables
library(knitr)
```

### Import Data
The data is the same data we used in the chapter on Ridge and Lasso Regression. Recall that we used data from the set for predict satellite survivability using multiple variables.

```{r}
library(glmnet)
train<-read.csv("C:\\Users\\jeff\\Documents\\Data\\Survive.csv") #Import the train set

train$RSO_Weight[is.na(train$RSO_Weight)] <- mean(train$RSO_Weight, na.rm = TRUE)
train$RSO_Density[is.na(train$RSO_Density)] <- "Low"
train$RSO_Visibility[train$RSO_Visibility == 0] <- mean(train$RSO_Visibility)
train$RSO_MRP[is.na(train$RSO_MRP)] <- mean(train$RSO_MRP, na.rm = TRUE)
train$Orbit_Establishment_Year=2013 - train$Orbit_Establishment_Year
train$Orbit_Height[is.na(train$Orbit_Height)] <- "LEO"
train$Stealth_Type[is.na(train$Stealth_Type)] <- "Stealth_1"
train$RSO_Type[is.na(train$RSO_Type)] <- "RSO_Type1"
```

### Drop Unecessary Variable
There are two binary identifier variable that we do not need and would not perform well in a regression model. They are better suited for classification.

```{r}
drop <- c("RSO_Identifier", "Orbit_Identifier")
train = train[,!(names(train) %in% drop)]
#head(train,5)
```

### Label Encoding
Here, we encode our categorical response variable into integer values (0,1), because all machine learning models require the data to be encoded into numerical format. LabelEncoder(), from the superml-package takes a vector of character or factor values and encodes them into numeric values.

```{r}
#library(superml)
lbl = LabelEncoder$new()
train$RSO_Type = lbl$fit_transform(train$RSO_Type)
train$RSO_Name = lbl$fit_transform(train$RSO_Name)
train$RSO_Density = lbl$fit_transform(train$RSO_Density)
train$Orbit_Height = lbl$fit_transform(train$Orbit_Height)
train$Stealth_Type = lbl$fit_transform(train$Stealth_Type)
#train$Orbit_Identifier = lbl$fit_transform(train$Orbit_Identifier)
#train$RSO_Identifier = lbl$fit_transform(train$RSO_Identifier)
```

### Normalizing
Normaliziing data allows us to analyze and model data that are of the same/similar scale, which makes model more accurate.

```{r}
train$RSO_Type = (train$RSO_Type-(mean(train$RSO_Type))/sd(train$RSO_Type))
train$RSO_Name = (train$RSO_Name-(mean(train$RSO_Name))/sd(train$RSO_Name))
train$RSO_Density = (train$RSO_Name-(mean(train$RSO_Density))/sd(train$RSO_Density))
#train$RSO_Type = (train$RSO_Type-(mean(train$RSO_Type))/sd(train$RSO_Type))
train$Stealth_Type = (train$Stealth_Type-(mean(train$Stealth_Type))/sd(train$Stealth_Type))
train$Orbit_Height = (train$Orbit_Height - (mean(train$Orbit_Height))/ sd(train$Orbit_Height))
#train$RSO_Identifier = (train$RSO_Identifier - (mean(train$RSO_Identifier))/ sd(train$RSO_Identifier))
#summary(train)
```

### Write and Read the Data

To preserve the transformations we made on the data, it is wise to store it to protect the original data. Then we'll read the same data back in our environment.

```{r}
#set.seed(222)
setwd("C:/Users/jeff/Documents/Data")
write_csv(train,"sun_data.csv")
surv_data<-read_csv("C:/Users/jeff/Documents/Data/surv2_data.csv")
head(surv_data,5)
```

## Remove Rows with NA
Even though we have preprocessed the data for miss values, there are probably missing values in our coolection.

```{r}
train <- train[rowSums(is.na(train)) == 0, ]
```

### Split the Data into Subsets 
Finally, we split the set into the training set and the cross-validation set, which will complete our data preprocessing.

```{r}
set.seed(567)
part <- sample(2, nrow(X), replace = TRUE, prob = c(0.7, 0.3))
X_train<- X[part == 1,]
X_cv<- X[part == 2,]

Y_train<- Y[part == 1,]
Y_cv<- Y[part == 2,]

```

### Initial Lambda
Here we assign an initial $\lambda$ value for both the ridge and lasso regression  models.

```{r}
lambda <- 10^seq(10, -2, length = 100)
```

### Ridge Regression Model

We start with the Ridge regression that we learned about in Chapter  

```{r}
ridge_reg <- cv.glmnet(X[X_train,], Y[X_train], alpha = 0, lambda = lambda)
#ridge_reg <- glmnet(X, Y, alpha = 0, lambda = lambda)
#find the best lambda via cross validation
bestlam <- ridge_reg$lambda.min
ridge.pred <- predict(ridge_reg, llambda = bestlam, newx = X)
ridge.cv <- predict(ridge_reg, llambda = bestlam, newx = X_cv)
```

### Ridge Model Performance
Next, we check the model performance that will use for comparison with our methods.

```{r}
SSE <- sum((ridge.pred - Y_train)^2)
SSE_val <- sum((ridge.cv - Y_cv)^2)
SSR <- sum((Y_train - mean(Y_train))^2)
SSR_val <- sum((Y_cv - mean(Y_cv))^2)
SST <- SSR + SSE
SST_val<-SSR_val+SSE_val
Ridge_R_sq <- SSR / SST
Ridge_RMSE = sqrt(SSE/nrow(X_train))
Ridge_val_R_sq <- SSR_val / SST_val
Ridge_val_RMSE = sqrt(SSE_val/nrow(X_cv))
print(paste("Ridge R-square =", Ridge_R_sq))
print(paste("Ridge Regression RMSE =", Ridge_RMSE))
print(paste("Ridge CV R-square =", Ridge_val_R_sq))
print(paste("Ridge CV RMSE =", Ridge_val_RMSE))
```

## Lasso Regression Model
Now, we fit a lasso regression model to the same sun azimuth data.

```{r}
lasso_reg <- glmnet(X[X_train,], Y[X_train], alpha = 1, lambda = lambda)
lasso.pred <- predict(lasso_reg, s = bestlam, newx = X[X_cv,])
SSE <- sum((lasso.pred - Y)^2)
SSR <- sum((Y[X_cv] - mean(Y[X_cv]))^2)
SST <- SSR + SSE
R_square <- SSR / SST
Lasso_RMSE = sqrt(SSE/nrow(Y))
print(paste("Lasso R-square =",R_square))
print(paste("Lasso Regression RMSE =", Lasso_RMSE))
print(paste("Ridge Regression RMSE =", Ridge_RMSE))
```

### Optimal Lambda
Here we assign an initial lambda value for the Ridge and Lasso models. The system, like `optimal_lambda`.

```{r}
optimal_lambda <- cv_lasso$lambda.min
glmnet(X[X_train,], Y[X_train], alpha = 1, family = 'gaussian', lambda = optimal_lambda, thresh = 1e-07)
lasso.pred <- predict(lasso_reg, s = optimal_lambda, newx = X)
lasso.cv <- predict(lasso_reg, s = optimal_lambda, newx = X_cv)
```

### Lasso Model Performance
Finally, we obtain the performance of the Lasso model for comparison with the other models.

```{r}
SSE <- sum((lasso.pred - Y_train)^2)
SSE_val <- sum((lasso.cv - Y_cv)^2)
SSR <- sum((Y_train - mean(Y_train))^2)
SSR_val <- sum((Y_cv - mean(Y_cv))^2)
SST <- SSR + SSE
SST_val<-SSR_val+SSE_val
Lasso_R_sq <- SSR / SST
Lasso_RMSE = sqrt(SSE/nrow(X_train))
Lasso_val_R_sq <- SSR_val / SST_val
Lasso_val_RMSE = sqrt(SSE_val/nrow(X_cv))
print(paste("Lasso R-square =", Lasso_R_sq))
print(paste("Lasso Regression RMSE =", Lasso_RMSE))
print(paste("Lasso CV R-square =", Lasso_val_R_sq))
print(paste("Lasso CV RMSE =", Lasso_val_RMSE))
```

## ElaticNet Regression Model
Without much fanfare, we expose the ElasticNet Regression, which`discuss in more detail in Part II of this chapter.

### Training and Testing Data Split
Here we perform a dataset split that well use at the very end of this section for the purpose of validating the model.

```{r}
ind <- sample(2, nrow(train), replace = TRUE, prob = c(0.7, 0.3))
train <- train[ind==1,]
test <- train[ind==2,]
```

### creating Custom Control Parameters
This "helper" function redusce the complecty in defining the ElasticNet model we will connstruct.

```{r}
control <- trainControl(method = "repeatedcv",
                              number = 5,
                              repeats = 5,
                              search = "random",
                              verboseIter = TRUE)
```

### Training ELastic Net Regression model

```{r,  results='hide'}
elastic_model <- train(Survivability ~ .,
                           data = train,
                           method = "glmnet",
                           #preProcess = c("center", "scale"),
                           tuneLength = 10,
                           trControl = control)
```

### Print Elastic Model Results

```{r}
elastic_model
```

### Find Best Values

```{r}
set.seed(123)

# Train Control
train.control <- trainControl(method = "repeatedcv", number = 10, repeats = 10 , search = "random")

# Train the Model
cv_for_best_value <- train(Survivability ~ ., data = train, method="glmnet", trControl = train.control)

# Obtaining Best Value of $\alpha$ and $\lambda$
cv_for_best_value$bestTune
```


# Alternative Performance Function
Here, we constructed a user defined function to compute $R^2$ and RMSE for true and predicted values

```{r}
eval_results <- function(true, predicted, df) {
  SSE <- sum((predicted - true)^2)
  SST <- sum((true - mean(true))^2)
  R_square <- 1 - SSE / SST
  RMSE = sqrt(SSE/nrow(df))
  # Model performance metrics
data.frame(
  RMSE = RMSE,
  Rsquare = R_square
)
}
```


### Implement Best Values

```{r}
sun_elnet <- glmnet(X, Y, alpha =cv_for_best_value$bestTune$alpha , lambda = cv_for_best_value$bestTune$lambda ,family = "gaussian")
sun_elnet
  
# Model Prediction
elnet_pred <- predict(cv_for_best_value, X)
eval<-eval_results(Y, elnet_pred, X)
  
# Multiple R-squared
rsq <- cor(X, elnet_pred)^2
rsq

### Plot ElasticNet Model
 
# Plot
plot(cv_for_best_value, pch=16, cex = 2, main = "Elastic Net Regression") 

```

### Model Performance

```{r}
library(Metrics)
library(rsq)
mean(cv_for_best_value$resample$RMSE)
mean(cv_for_best_value$resample$Rsquared)
mean(cv_for_best_value$resample$MAE)
```

### Plot Variable Importance

```{r}
plot(varImp(cv_for_best_value), pch=16, lwd = 2, cex = 2)
```

### Use Alternative Performance Function

```{r}
Elnet_RMSE <- eval$RMSE
Elnet_R_sq <- eval$Rsquare

tab_01 = data.frame(
  Measure = c("Lasso RMSE", "Lasso R.Sq", "Ridge RMSE", "Ridge R.Sq", "Elnet RMSE", "Elnet R.Sq"),
  Value  = c(Lasso_RMSE, Lasso_R_sq, Ridge_RMSE, Ridge_R_sq, Elnet_RMSE, Elnet_R_sq)
)

knitr::kable(tab_01, digits = 6)
```

### Performance Table

```{r}
A<-rbind(Lasso_R_sq, Ridge_R_sq, Elnet_R_sq) %>% round(7)
B<-rbind(Lasso_RMSE, Ridge_RMSE, Elnet_RMSE) %>% round(7)
D<-cbind(A, B)
C<-cbind("R.Square", "RMSE")
rbind(C,D)
```

So, how did we get a negative $R^2$ value? In short, $R^2$ is only the square of correlation if we happen to be (1) using linear regression models, and (2) are evaluating them on the same data they are fitted. 

# Compute R^2 from true and predicted values

```{r}
elnet_pred <- predict(cv_for_best_value, testing)
eval<-eval_results(testing$Survivability, elnet_pred, testing)
eval
```


```{r}
mod1<-glmnet(X, Y, alpha=0, family = "gaussian")
print(paste("Variable: Intercept =", mod1.coef[1]))
print(paste("Variable:", names(data01), "=", 
mod1.coef[2:43]))
barplot(mod1.coef, main = "Model 1 Coefficients", ylab = "Coefficients", las = 2)
plot(mod1, lwd = 3)
```


```{r}
mod2<-cv.glmnet(X, Y, alpha=0, nfolds=5, type.measure = "mse")
summary(mod2)
mod2.coef <- predict(mod1, type = "coefficients", s = bestlam)[1:43]
print(paste("Variable: Intercept =", mod2.coef[1]))
print(paste("Variable:", names(data01), "=", 
mod2.coef[2:43]))
barplot(mod2.coef, main = "Model 2 Coefficients", ylab = "Coefficients", las = 2)
plot(mod2, lwd = 2)
```
