---
title: "Orbital Parameters"
author: "Jeffrey Strickland"
date: '2022-09-06'
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, dpi = 300)
```

#The ECI coordinates can then be calculated. In this case, they will be in TEME frame, since the original orbital elements are derived from a TLE

```{r}
library(asteRisk)
#try(coordinates_ECI <- KOEtoECI(a0, e0, i0, M0, omega0, OMEGA0))
```


#Let's calculate the ECI coordinates from the orbital elements provided by a TLE. It should be noted that this is often not recommended, since the orbital elements supplied in a TLE are not osculating orbital elements, but instead mean orbital elements set to fit a range of actual observations. The recommended procedures are to use TLE only in conjunction with the SGP4/SDP4 models, and viceversa.


```{r}
test_TLEs <-readTLE(paste0(path.package("asteRisk"),"//testTLE.txt"))
#italsat2_TLE <- parseTLElines(lines=3)
#italsat2_TLE
```


```{r}
library(HH)
```

#The following orbital parameters correspond to an object with NORAD catalogue number 24208 (Italsat 2) the 26th of June, 2006 at 00:58:29.34 UTC

# The period is lower than 225 min, and therefore the SGP4 model is valid.
# Lets calculate the position and velocity of the satellite 40 minutes after
# epoch


```{r}
set.seed(42)
g = 6.67*10^(-11)
gen_some_data = function(n_obs = 50) {
  mu = 3.986004418e14 # gravitational parameter
  int_designator = seq(0, 10, length.out = n_obs)
  eccentricity = runif(n = n_obs, min = 0, max = 2.5)
  focal_param = round(runif(n = n_obs, min = 100, max = 400),4)
  semi_latus_rectum = focal_param*eccentricity  
  semimajor_axis = semi_latus_rectum/abs((1-eccentricity^2))
  classification = sample(c("U", "C", "S"), size = n_obs, replace = TRUE)
  periapsis  = round(runif(n = n_obs, min = 0, max = 360), 4)
  inclination = round(runif(n = n_obs, min = 0, max = 180), 4)
  mean_motion = 360/focal_param
  mean_anomoly <- 2*pi/runif(n = n_obs, min = 0, max = 24) 
  pos_y <- round(runif(n = n_obs, min = -1000, max = 1000),4)
  pos_z <- round(runif(n = n_obs, min = -1000, max = 1000),4)
  omega0 <- atan2(pos_y,pos_z) 
  OMEGA0 <- 80.0121*pi/round(runif(n = n_obs, min = 180, max = 360), 4)  
  nu = round(runif(n = n_obs, min = 0, max = 360), 4)
  longitudePerigee <- OMEGA0 + omega0
  trueLongitude <- longitudePerigee + nu
  #mean_anomoly = round(runif(n = n_obs, min = 0, max = 360),4)
  day = round(runif(n = n_obs, min = 0, max = 24), 4)
  Bstar <- 0.66816e-4 # drag coefficient
  data.frame(mean_motion, semimajor_axis, semi_latus_rectum, eccentricity, periapsis, mean_anomoly, inclination, omega0, OMEGA0, trueLongitude, longitudePerigee)
}
```


```{r}
set.seed(42)
orb_data = gen_some_data(n_obs = 2500)
orb_idx  = sample(1:nrow(orb_data), 500)
orb_trn  = orb_data[orb_idx, ]
orb_val  = orb_data[-orb_idx, ]
summary(orb_trn)
nrow(orb_trn)
nrow(orb_val)
```


# Calculation of the orbital period
#2*pi/n0

```{r}
set.seed
orb_data = gen_some_data(n_obs = 5000)
orb_tst = gen_some_data(n_obs = 800)
orb_trn = gen_some_data(n_obs = 400)
summary(orb_trn)
```

## Test

```{r}
orb_trn <- as.data.frame(orb_trn)
head(orb_trn, 10)
```


# The period is lower than 225 min, and therefore the SGP4 model is valid.
# Lets calculate the position and velocity of the satellite 40 minutes after
# epoch


```{r}
new_lines <- cbind(n0=orb_trn$n0, e0=orb_trn$e0, i0=orb_trn$i0, M0=110.5714*pi/180, omega0=orb_trn$omega0, OMEGA0=orb_trn$OMEGA0, Bstar=orb_trn$Bstar)

```

## SGP4 Propogator

```{r}
new_state1 <- sgp4(n0=(orb_trn[1,1]), e0=0.0086731, i0=orb_trn[1,3], M0=orb_trn[1,4],omega0=orb_trn[1,5],OMEGA0=orb_trn[1,6],Bstar=orb_trn[1,7], targetTime = 40)
new_state1
```

## Experiment 1

```{r}
#new_state[1:i]
new_state0 <- sgp4(n0=100, e0=0.0086731, i0=200, M0=300,omega0=400,OMEGA0=500,Bstar=600, targetTime = 20)
print(new_state0)
```

## Experiment 2

```{r}
XX<-sgp4(n0=100, e0=0.0086731, i0=200, M0=300,omega0=400,OMEGA0=500,Bstar=600, targetTime = 20)
print(XX)
```

## Test

```{r}
for (val in 1: 28){ 
new_lines[val] <- sgp4(n0=((2*pi)/(1440)), 
                  e0=test_TLEs[[val]]$eccentricity, 
                  i0=test_TLEs[[val]]$inclination, 
                  M0=test_TLEs[[val]]$meanAnomaly, 
                  omega0=test_TLEs[[val]]$perigeeArgument, 
                  OMEGA0=test_TLEs[[val]]$ascension,                 #trueLongitude=test+TLEs[[val]]$trueLongitude,                  #longitudePerigee=test+TLEs[[val]]$longitudePerigee,
                  Bstar=test_TLEs[[val]]$Bstar,targetTime = 20)
}
```


```{r}
# The following orbital parameters correspond to an object with NORAD catalogue.
# number 88888 the 1st of October, 1980 at 23:41:24 UTC.
g = 6.67*10^(-11)

gen_orb_data = function(n_obs = 100) {
  pos_y <- round(runif(n = n_obs, min = -1000, max = 1000),4)
  pos_z <- round(runif(n = n_obs, min = -1000, max = 1000),4)
  int_designator = seq(0, 10, length.out = n_obs)
  focal_param = round(runif(n = n_obs, min = 100, 
max = 400),4)
  classification = sample(c("U", "C", "S"), size = n_obs, 
replace = TRUE)
  eccentricity = runif(n = n_obs, min = 0, max = 90)
  periapsis  = round(runif(n = n_obs, min = 0, max = 90), 4)
  inclination = round(runif(n = n_obs, min = 0, max = 180), 4)
  semi_latus_rectum = focal_param*eccentricity
  semimajor_axis = semi_latus_rectum/abs((1-eccentricity^2))
  mean_motion = 360/focal_param
  mean_anomoly <- 2*pi/runif(n = n_obs, min = 0, max = 24) 
  omega0 <- atan2(pos_y, pos_z) 
  Bstar <- 0.66816e-4 # drag coefficient.

  data.frame(mean_motion, semimajor_axis, semi_latus_rectum, 
eccentricity, periapsis, mean_anomoly, inclination, 
focal_param, periapsis, omega0)
}
```


```{r}
calc_rmse = function(actual, predicted) {
  sqrt(mean((actual - predicted) ^ 2))
}
```



```{r}
num_sims = 300
num_degrees = 10
val_rmse = matrix(0, ncol = num_degrees, nrow = num_sims)
```



```{r}
set.seed(42)
for (i in 1:num_sims) {
  # simulate data.
  orb_data = gen_orb_data(n_obs = 300)
  # set aside validation set.
  orb_idx = sample(1:nrow(orb_data), 160)
  orb_trn = orb_data[orb_idx, ]
  orb_val = orb_data[-orb_idx, ]
  # fit models and store RMSEs.
  for (j in 1:num_degrees) {
    #fit model.
    fit = glm(mean_motion ~ poly(semi_latus_rectum, degree = j), data = orb_trn)
    # calculate error.
    val_rmse[i, j] = calc_rmse(actual = orb_val$mean_motion, predicted = predict(fit, orb_val))
  }
}

```

## Loop in Development

```{r}
set.seed(42)
orb_data = gen_orb_data(n_obs = 1000)
orb_idx  = sample(1:nrow(orb_data), 500)
orb_trn  = orb_data[orb_idx, ]
orb_val  = orb_data[-orb_idx, ]
summary(orb_trn)

```
Plot the mean motion vs semilatus rectum

```{r, echo = TRUE, fig.height = 4, fig.width = 8}
plot(mean_motion ~ semi_latus_rectum, data = orb_trn, 
col = "dodgerblue", pch = 20)
grid()
curve(x ^ 3, add = TRUE, col = "black", lwd = 2)

```

Fit mean motion using a 10 degree polynomial of the semilatus rectum.

```{r}
fit = lm(mean_motion ~ poly(semi_latus_rectum, 10), 
data = orb_trn)
summary(fit)
```

Now, we calculate the RMSE for the training and validation sets.

```{r}
calc_rmse(actual = orb_trn$mean_motion, predicted = predict(fit, orb_trn))
calc_rmse(actual = orb_val$mean_motion, predicted = predict(fit, orb_val))

```

Set up parameters for simulating a fit of mean motion iterating though 1-10 degree polynomial of the semilatus rectum.

```{r}
num_sims = 300
num_degrees = 10
val_rmse = matrix(0, ncol = num_degrees, nrow = num_sims)
```

Run the simulation and calculate the RMSE.

```{r}
set.seed(42)
for (i in 1:num_sims) {
  # simulate data.
  orb_data = gen_orb_data(n_obs = 300)
  # set aside validation set.
  orb_idx = sample(1:nrow(orb_data), 160)
  orb_trn = orb_data[orb_idx, ]
  orb_val = orb_data[-orb_idx, ]
  # fit models and store RMSEs.
  for (j in 1:num_degrees) {
    #fit model.
    fit = glm(mean_motion ~ poly(semi_latus_rectum, degree = j), data = orb_trn)
    # calculate error.
    val_rmse[i, j] = calc_rmse(actual = orb_val$mean_motion, predicted = predict(fit, orb_val))
  }
}
summary(fit)
```

Next, we plot the RMSE vs Degree of polynomial from the simulation and the model chosen vs degree of polynomial.

```{r, echo = TRUE, fig.height = 5, fig.width = 8}
par(mfrow = c(1, 2))
matplot(t(val_rmse)[, 1:10], pch = 20, type = "b", ylim = c(0.50, 0.75), xlab = "Polynomial Degree", ylab = "RMSE", main = "RMSE vs Degree")
barcol = c("grey", "grey", "dodgerblue", "grey", "grey", "grey", "grey", "grey", "grey", "grey")
barplot(table(factor(apply(val_rmse, 1, which.min), levels = 1:10)),
        ylab = "Times Chosen", xlab = "Polynomial Degree", col = barcol, main = "Model Chosen vs Degree")
```

Run the simulation using a 5-fold cross-validation set.

```{r}
cv_rmse = matrix(0, ncol = num_degrees, nrow = num_sims)
set.seed(42)
for (i in 1:num_sims) {
  # simulate data, use all data for training.
  orb_trn = gen_orb_data(n_obs = 200)
  # fit models and store RMSE.
  for (j in 1:num_degrees) {
    #fit model.
    fit = glm(mean_motion ~ poly(semi_latus_rectum, 
degree = j), data = orb_trn)
    # calculate error.
    cv_rmse[i, j] = sqrt(boot::cv.glm(orb_trn, fit, 
K = 5)$delta[1])
    
  }
}

```

Create a table of RMSE and 5-Fold CV RMSE with SDs.

```{r, echo = TRUE}
results = data.frame(
  degree = 1:10,
  colMeans(val_rmse),
  apply(val_rmse, 2, sd),
  colMeans(cv_rmse),
  apply(cv_rmse, 2, sd)
)
colnames(results) = c(
  "Polynomial Degree",
  "Mean, Val",
  "SD, Val",
  "Mean, CV",
  "SD, CV"
)

knitr::kable(results, digits = 3)
```

Summarize the fit.

```{r}
summary(fit)
```

Create a dataframe of the 5-Fold CV RMSEs by polynomial degrees.

```{r}
df<-as.data.frame(cbind(cbind(cv_rmse[1,1:10]), cbind(cv_rmse[2,1:10]), cbind(cv_rmse[3,1:10]), cbind(cv_rmse[4,1:10]), cbind(cv_rmse[5,1:10]), cbind(cv_rmse[6,1:10]), cbind(cv_rmse[7,1:10]), cbind(cv_rmse[8,1:10]), cbind(cv_rmse[9,1:10]), cbind(cv_rmse[10,1:10])))
df
```

Plot the dataframe.

```{r, echo = TRUE, fig.height = 6, fig.width = 10}
plot(df)
```

Rename the columns of the dataframe.

```{r}
colnames(df) <-c("Poly1", "Poly2", "Poly3", "Poly4", "Poly5", "Poly6", "Poly", "Poly8", "Poly9", "Poly10")
```

Create a datframe the same dataframe with and index.

```{r}
dfx<-as.data.frame(cbind(cbind(seq(1,10,1)),cbind(cv_rmse[1,1:10]), cbind(cv_rmse[2,1:10]), cbind(cv_rmse[3,1:10]), cbind(cv_rmse[4,1:10]), cbind(cv_rmse[5,1:10]), cbind(cv_rmse[6,1:10]), cbind(cv_rmse[7,1:10]), cbind(cv_rmse[8,1:10]), cbind(cv_rmse[9,1:10]), cbind(cv_rmse[10,1:10])))
dfx
```

Create a barplot of the mean RMSEs for the best five polynomial degrees.

```{r, echo = TRUE, fig.height = 5, fig.width = 8}
barplot(cbind(mean(dfx$V2),mean(dfx$V3),mean(dfx$V4),mean(dfx$V5),mean(dfx$V6),mean(dfx$V7),mean(dfx$V8),mean(dfx$V9),mean(dfx$V10)), col = "blue")
```

Build a connected dotplot of the top five 5-fold polynomials.

```{r, echo = TRUE, fig.height = 5, fig.width = 8}
library(dplyr)
library(ggplot2)
df <- as.data.frame(cbind(cbind(cv_rmse[1,1:10]), cbind(cv_rmse[2,1:10]), cbind(cv_rmse[3,1:10]), cbind(cv_rmse[4,1:10]), cbind(cv_rmse[5,1:10]), cbind(cv_rmse[6,1:10]), cbind(cv_rmse[7,1:10]), cbind(cv_rmse[8,1:10]), cbind(cv_rmse[9,1:10]), cbind(cv_rmse[10,1:10])))

colnames(df) <- c("Poly1", "Poly2", "Poly3", "Poly4", "Poly5", "Poly6", "Poly7", "Poly8", "Poly9", "Poly10")
 
df1 <- as.data.frame(df$Poly1)
df2 <- as.data.frame(df$Poly2)
df5 <- as.data.frame(df$Poly5)
df6 <- as.data.frame(df$Poly6)
df9 <- as.data.frame(df$Poly9)

g <- ggplot() +
  geom_point(aes(data=df1, y = df$Poly1, x = seq(1,10,1), 
color = "Poly1"), size = 3) +
  geom_point(aes(data=df2, y = df$Poly2, x = seq(1,10,1), 
color = "Poly2"), size = 3) +
  geom_point(aes(data=df5, y = df$Poly5, x = seq(1,10,1), 
color = "Poly5"), size = 3) +
  geom_point(aes(data=df6, y = df$Poly6, x = seq(1,10,1), 
color = "Poly6"), size = 3) +
  geom_point(aes(data=df9, y = df$Poly9, x = seq(1,10,1), 
color = "Poly9"), size = 3) +
  geom_line(aes(data= df1, y = df$Poly1, x = seq(1,10,1), 
color = "Poly1"), size = 1.5) +
  geom_line(aes(data= df2, y = df$Poly2, x = seq(1,10,1),
color = "Poly2"), size = 1.5) +
  geom_line(aes(data= df5, y = df$Poly5, x = seq(1,10,1),
color = "Poly5"), size = 1.5) +
  geom_line(aes(data= df6, y = df$Poly6, x = seq(1,10,1),
color = "Poly6"), size = 1.5) +
  geom_line(aes(data= df9, y = df$Poly9, x = seq(1,10,1),
color = "Poly9"), size = 1.5) +
  labs(x = "Simulation Number") +
  labs(y = "CV RMSE by Poly") + 
  theme_classic()
g
```

Generate a new dataframe for plotting with a stacked barplot.

```{r}
dfz<-as.data.frame(cbind(cbind(cv_rmse[1,1:10]), cbind(cv_rmse[2,1:10]), cbind(cv_rmse[3,1:10]), cbind(cv_rmse[4,1:10]), cbind(cv_rmse[5,1:10]), cbind(cv_rmse[6,1:10]), cbind(cv_rmse[7,1:10]), cbind(cv_rmse[8,1:10]), cbind(cv_rmse[9,1:10]), cbind(cv_rmse[10,1:10])))
dfz
colnames(dfz) <-c("Poly1", "Poly2", "Poly3", "Poly4", "Poly5", "Poly6", "Poly7", "Poly8", "Poly9", "Poly10")
```

Creat the stacked barplot.

```{r, echo = TRUE, fig.height = 5, fig.width = 8}
mp <- barplot(as.matrix(dfz),col="blue") # default
tot <- colSums(dfz)
title(main = "5-Fold Cross Validation", font.main = 4)
```

Create an unstacked barplot with RMSE values by deggree of the polynomials.

```{r, echo = TRUE, fig.height = 5, fig.width = 8}
tot <- colMeans(dfz)
barplot(as.matrix(dfz), beside = TRUE, ylim = c(0, 5),col = c("lightblue", "mistyrose", "lightcyan", "lavender", "cornsilk", "dodgerblue", "violet", "red2", "orange", "green"))
title(main = "5-Fold Cross Validation", font.main = 4)
```

The function `which.min` determines the location, i.e., index of the (first) minimum of a numeric vector. levels provides access to the levels attribute of a variable. In `max_correct`, the argument `table(apply(val_rmse, 1, which.min))` in `max_correct` generates a table of bar heights and the maximum value in the table is 31, wile the maximum value of `table(apply(cv_rmse, 1, which.min))` is 36. These constitute the y-axes.

```{r, echo = TRUE, fig.height = 5, fig.width = 8}
par(mfrow = c(1, 2))
max_correct = max(max(table(apply(val_rmse, 1, which.min))), max(table(apply(cv_rmse, 1, which.min))))
barcol = c("grey", "grey", "dodgerblue", "grey", "grey", "grey", "grey", "grey", "grey", "grey")
barplot(table(factor(apply(val_rmse, 1, which.min), levels = 1:10)), ylim = c(0, max_correct), 
        ylab = "Times Chosen", xlab = "Polynomial Degree", col = barcol, main = "Single Validation Set")
barplot(table(factor(apply(cv_rmse,  1, which.min), levels = 1:10)), ylim = c(0, max_correct), 
        ylab = "Times Chosen", xlab = "Polynomial Degree", col = barcol, main = "5-Fold Cross-Validation")
```

Single validation set vs 5-Fold cross-validation.

```{r, echo = TRUE, fig.height = 5, fig.width = 8}
par(mfrow = c(1, 2))
matplot(t(val_rmse)[, 1:10], pch = 20, type = "b", ylim = c(0.50, 0.80), xlab = "Polynomial Degree", ylab = "RMSE", main = "Single Validation Set")
matplot(t(cv_rmse)[, 1:10],  pch = 20, type = "b", ylim = c(0.50, 0.80), xlab = "Polynomial Degree", ylab = "RMSE", main = "5-Fold Cross-Validation")
```

Backup table

```{r, echo = FALSE}
print(paste("Poly Degrees  |Mean, Val| SD, Val |Mean, CV | SD, CV"))
print(paste("-----------------------------------------------------"))
print(paste(" 1            |",round(mean(val_rmse[,1]),5),"|",round(sd(val_rmse[,1]),5)," |",round(mean(cv_rmse[,1]),5),"|",round(sd(cv_rmse[,1]),5)))
print(paste(" 2            |",round(mean(val_rmse[,2]),5),"|",round(sd(val_rmse[,2]),5),"|",round(mean(cv_rmse[,2]),5),"|",round(sd(cv_rmse[,2]),5)))
print(paste(" 3            |",round(mean(val_rmse[,3]),5),"|",round(sd(val_rmse[,3]),5),"|",round(mean(cv_rmse[,3]),5),"|",round(sd(cv_rmse[,3]),5)))
print(paste(" 4            |",round(mean(val_rmse[,4]),5),"|",round(sd(val_rmse[,4]),5),"|",round(mean(cv_rmse[,4]),5),"|",round(sd(cv_rmse[,4]),5)))
print(paste(" 5            |",round(mean(val_rmse[,5]),5),"|",round(sd(val_rmse[,5]),5),"| ",round(mean(cv_rmse[,5]),5),"|",round(sd(cv_rmse[,5]),5)))
print(paste(" 6            |",round(mean(val_rmse[,6]),5),"|",round(sd(val_rmse[,6]),5),"|",round(mean(cv_rmse[,6]),5),"|",round(sd(cv_rmse[,6]),5)))
print(paste(" 7            |",round(mean(val_rmse[,7]),5),"|",round(sd(val_rmse[,7]),5),"|",round(mean(cv_rmse[,7]),5),"|",round(sd(cv_rmse[,7]),5)))
print(paste(" 8            |",round(mean(val_rmse[,8]),5),"|",round(sd(val_rmse[,8]),5),"| ",round(mean(cv_rmse[,8]),5),"|",round(sd(cv_rmse[,8]),5)))
print(paste(" 9            |",round(mean(val_rmse[,9]),5),"|",round(sd(val_rmse[,9]),5),"|",round(mean(cv_rmse[,9]),5),"|",round(sd(cv_rmse[,9]),5)))
print(paste("10            |",round(mean(val_rmse[,10]),5),"|",round(sd(val_rmse[,10]),5),"| ",round(mean(cv_rmse[,10]),5),"|",round(sd(cv_rmse[,10]),5)))

```

